---
title: "Stack Barplot for taxonomy"
author: "Sebastián Sáenz"
date: "30/10/2020"
output: github_document
---
We are going to use **ggplot** to create our barplot. Load the library **tidyverse** 
or install it in case you have not done it before.
```{r, warning=FALSE}
library(tidyverse)
```

Next we are going to load our data. The relative abundance of the different taxa 
were previously calculated. **check.names** prevent that R would change
your column names. Use **str()** to get information about your data frame. In this case
we have 17 rows and 8 columns. **str()** also show you what kind of variable you have.
```{r load your data}

df <- read.table("mock_tax.txt",
                 header = TRUE,
                 sep = "\t",
                 check.names = FALSE)

str(df)


colSums(df[3:8])

```
Next we transform our df to a longer format as it is easier to use with **ggplot**.
We create two new variables **Treatment** and **Abundance**. Treatment would receive all 
the column names and Abundance all the values.

```{r transform df}
df_long <- pivot_longer(df, cols = c(3:4),
             names_to = "Treatment",
             values_to ="Abundance")

str(df_long)
```
Now we have 102 rows and 4 columns.


```{r }
relative_abundace <- df_long %>%
    group_by(Family, Treatment) %>%
    summarise(relative_abundance = sum(Abundance)/3)
  
```

## Ploting

```{r }
sort.class <- relative_abundace %>% 
  count(Family, wt = relative_abundance) %>%
  arrange(desc(n)) %>%
  pull(Family) 

relative_abundace %>%
count(Family, wt =relative_abundance)




relative_abundace %>%
  mutate(Family = factor(Family, levels = sort.class)) %>%
  ggplot(aes(x = Treatment, 
             y = relative_abundance, 
             fill = Family)) +
  geom_bar(stat = 'identity',
           position = position_dodge()) +
  scale_fill_brewer(palette = "Dark2") +
  facet_wrap(~Treatment) +
  theme_minimal()


col_list <- c('#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99','#e31a1c','#fdbf6f','#ff7f00')

relative_abundace %>%
  ggplot(aes(x = Treatment, y = relative_abundance, fill = Family)) +
  geom_bar(stat = 'identity') +
  scale_fill_manual(values = col_list)
```

